\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{siunitx}
\usepackage{pgfplots}
\usepackage{url}
\usepackage{hyperref}

\pgfplotsset{compat=1.18}
\sisetup{round-mode=places,round-precision=3}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{Latency, Not FLOPs, Predicts Inference Energy on Apple Silicon: \\
A Five-Architecture Benchmark with Residualized GAN Trial Modeling}

\author{\IEEEauthorblockN{Aryan Shah}
\IEEEauthorblockA{\textit{Texas Academy of Mathematics and Science} \\
\textit{University of North Texas}\\
Denton, TX, USA \\
aryanshah.2431@gmail.com}}

\maketitle

\begin{abstract}
We study which neural network architectures waste the most energy during
CPU-only inference on Apple Silicon, and we develop a paper-aligned synthetic
trial modeling pipeline to support repeated-trial analysis when measurement
budgets are limited. Across five architectures (MobileNetV3-Small, MobileNetV2,
ResNet-18, Tiny-ViT-5M, EfficientNet-B0), we find that latency predicts
measured energy nearly perfectly ($R^2 = 0.998$), while FLOPs is a substantially
weaker predictor ($R^2 = 0.62$). EfficientNet-B0 consumes 2.65$\times$ the
energy of ResNet-18 despite having 4.7$\times$ fewer FLOPs, confirming on
modern Apple Silicon that architectural complexity proxies do not capture
deployment energy. We further report Energy-Delay Product (EDP) as a unified
efficiency metric, under which MobileNetV3-Small is 13.4$\times$ more efficient
than EfficientNet-B0. To support repeated-trial analysis without re-running the
full benchmark, we develop a residualized conditional WGAN-GP. An initial
configuration trained on raw physical values across heterogeneous devices
exhibited mode collapse; switching to combo-normalized residuals within a
single-device scope resolved this. The revised GAN reproduces measured means
within 1.0--2.3\% on energy and preserves the measured ranking. We frame the
synthetic spreads as support estimates for repeated trials and identify direct
validation against independent measurements as required future work. The
methodological lesson generalizes: synthetic models of hardware behavior under
sparse data must be tightly scoped and trained on normalized residual structure
rather than raw heterogeneous values.
\end{abstract}

\begin{IEEEkeywords}
edge AI, energy efficiency, neural network inference, generative adversarial
networks, Apple Silicon, benchmarking, energy-delay product
\end{IEEEkeywords}

\section{Introduction}
As neural networks are increasingly deployed on mobile, embedded, and
edge-class hardware, energy efficiency has become a first-order deployment
constraint. Traditional architecture metrics such as floating-point operations
(FLOPs), parameter count, and model size are widely used as surrogates for
efficiency, but these proxies do not reliably predict real hardware energy
expenditure \cite{horowitz2014,mittal2019,hardwareaware2019,eyeriss2017}.

This paper makes three contributions:

\textbf{(1) An empirical benchmark on Apple Silicon} measuring inference
energy for five neural network architectures under a unified CPU-only PyTorch
runtime. We quantify how strongly common proxies predict measured behavior and
report Energy-Delay Product as a unified efficiency metric. The central finding
is that latency, not FLOPs, predicts measured energy on this platform.

\textbf{(2) An interpretive analysis} of why specific architectural choices
(depthwise convolutions, transformer attention, compound scaling) interact with
the CPU memory hierarchy and execution backend in ways that decouple FLOPs from
measured energy.

\textbf{(3) A residualized GAN methodology} for synthetic trial modeling under
sparse-data conditions. We document a mode collapse failure under raw-scale,
multi-device training and show that combo-normalized residuals within a
single-device scope resolve it. We are explicit that the GAN's fidelity is
evaluated against grounded training distributions rather than independent
held-out trials, and we frame the synthetic spreads as support estimates
rather than validated predictions.

The synthetic component is motivated by a gap in measurement-campaign
budgets. Re-running large benchmark sweeps with full trial-level logging is
expensive on edge hardware due to thermal management, sampling overhead, and
runtime-state isolation. A scope-matched synthetic model that preserves
measured means and rankings is useful for sensitivity studies and ablations
that would otherwise require additional hardware time.

\section{Related Work}
Research on efficient neural network deployment spans three relevant threads.
The first focuses on architecture design, including MobileNet, MobileNetV2,
MobileNetV3, and EfficientNet, which reduce computation through depthwise
separable convolutions, inverted residuals, and compound scaling
\cite{mobilenet2017,mobilenetv22018,mobilenetv32019,efficientnet2019}. The
second studies systems-level energy behavior on real hardware, emphasizing
that data movement, memory access, and execution backends often dominate the
cost captured by abstract FLOP counts
\cite{horowitz2014,mittal2019,eyeriss2017,hardwareaware2019}. The third
concerns synthetic or predictive modeling of hardware behavior, where
statistical or machine learning models substitute for full measurement
campaigns when those are impractical.

This paper sits primarily in the second thread, with a methodological
contribution to the third. The primary evidence is a measurement study on
Apple Silicon. The synthetic extension models repeated-trial behavior under
the same device and runtime scope and contributes a process lesson on
synthetic-model scope sensitivity.

\section{Methodology}

\subsection{Measured Inference Benchmark}
All benchmark runs used CPU-only inference in PyTorch with a fixed input
tensor of size $1 \times 3 \times 224 \times 224$. Each model was executed
with warm-up iterations followed by repeated timed inference runs.

CPU-only inference was selected for two reasons. First, it isolates a runtime
configuration in which deployment-time energy is governed by the general-purpose
core and memory hierarchy rather than a specialized accelerator, which is the
common case for cross-platform edge frameworks that lack hardware-specific
backends for Apple Silicon. Second, it removes accelerator scheduling
nondeterminism (Neural Engine, Metal Performance Shaders) that would otherwise
confound model-to-model comparison under a unified runtime.

Inference was performed locally with background activity minimized. To reduce
thermal-throttling effects on Apple Silicon, model order was randomized across
runs and a fixed cooldown period was inserted between models. Power was sampled
using \texttt{powermetrics} at 1 Hz. Because individual inference latencies
(15--55 ms) are shorter than the sampling interval, energy was attributed by
running many consecutive inferences within a measurement window and dividing
the integrated window energy by the number of inferences. This window-level
attribution is reported transparently as a methodological detail rather than
treated as per-inference power resolution.

\begin{table}[ht]
\centering
\caption{Hardware and software platform.}
\begin{tabular}{ll}
\toprule
Component & Specification \\
\midrule
Device & Apple MacBook Pro (Apple Silicon) \\
Execution mode & CPU-only inference \\
Runtime & PyTorch (version logged at runtime) \\
Operating system & macOS 26.2 \\
Memory & 8 GB RAM \\
Power measurement & \texttt{powermetrics}, 1 Hz \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Models Evaluated}
The benchmark covers five architectures spanning depthwise CNNs, a baseline
CNN, and a lightweight transformer.

\begin{table}[ht]
\centering
\caption{Architectures evaluated.}
\begin{tabular}{l l S[table-format=2.2] S[table-format=1.2]}
\toprule
Model & Type & {Params (M)} & {FLOPs (G)} \\
\midrule
MobileNetV3-Small & CNN (depthwise) & 2.54 & 0.22 \\
MobileNetV2 & CNN (depthwise) & 3.50 & 0.30 \\
ResNet-18 & CNN (baseline) & 11.69 & 1.82 \\
Tiny-ViT-5M & Transformer & 12.08 & 1.30 \\
EfficientNet-B0 & CNN (scaled) & 5.29 & 0.39 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Measured Metrics and Energy-Delay Product}
We report energy per inference $E$ (J), latency per inference $L$ (ms), and
average power $P$ (W). Let $P(t)$ denote sampled power over an inference window
$[t_0, t_1]$. Energy is approximated by
\begin{equation}
E \approx \int_{t_0}^{t_1} P(t)\,dt \approx \sum_{k=1}^{K} P_k \Delta t.
\end{equation}
Average power is reported as $P = E \cdot 1000 / L$, computed per model from
each model's energy and latency entries. This avoids reporting a flat aggregate
power as if it carried model-specific precision.

We additionally report Energy-Delay Product (EDP), a standard metric in
low-power computing \cite{horowitz2014} that combines energy and latency into
a single efficiency score:
\begin{equation}
\text{EDP} = E \cdot L.
\end{equation}
Lower EDP indicates a model that is both faster and more energy-efficient. EDP
is reported in units of J$\cdot$ms.

\subsection{Paper-Aligned Synthetic Trial Modeling}
The synthetic extension is deliberately narrow. The synthetic pipeline uses
only the Apple-Silicon device, the same five models, and observed-only
training combinations.

The paper-aligned benchmark was stored as one measured row per model. Each
measured row was expanded into a grounded seed set by sampling around the
reported means with conservative noise levels. These grounded samples served
as the empirical training distribution for a conditional WGAN-GP.

\subsection{Mode Collapse Under Raw-Scale Multi-Device Training}
An initial GAN configuration trained directly on raw physical values
(\texttt{energy\_J}, \texttt{power\_W}, \texttt{latency\_ms},
\texttt{energy\_std}) across a multi-device dataset (Apple Silicon, Raspberry
Pi, Jetson Nano, Coral Edge TPU, STM32) exhibited mode collapse. The dynamic
range across devices was too large: energy values spanned roughly two orders
of magnitude, latencies spanned three, and the per-device variance dominated
the within-model variance. The generator collapsed toward a single average
output that did not preserve per-model differences within any device.

\subsection{Combo-Normalized Residual Representation}
The final design uses combo-normalized residuals. For each
$(\text{device}, \text{model})$ pair $c$, we compute the per-combo mean
$\mu_c$ and standard deviation $\sigma_c$ for each continuous feature. Each
training sample $x$ is converted to a clipped $z$-score:
\begin{equation}
\tilde{x} = \mathrm{clip}\!\left(\frac{x - \mu_c}{\sigma_c}, -3, 3\right).
\end{equation}
The GAN learns a distribution over $\tilde{x}$ rather than $x$. During
generation, samples are mapped back to physical units using the combo-specific
$\mu_c$ and $\sigma_c$. This forces the GAN to learn within-combo structure
rather than relearning the larger between-combo differences.

In the Apple-only deployment used for the final paper, there are five observed
combinations and no estimated cross-device rows. The combo-normalized
representation is therefore equivalent to per-model normalization, which keeps
the resulting model tightly aligned with the paper's scope.

\subsection{Validation Scope of the Synthetic Pipeline}
We are explicit about what the synthetic pipeline can and cannot establish.
The GAN is trained on grounded samples drawn around the measured means. Its
fidelity evaluation, reported in Section~\ref{sec:fidelity}, therefore
demonstrates that the GAN reproduces the structure of its training
distribution rather than the structure of an independent set of held-out
measured trials. We frame the synthetic spreads as support estimates for
repeated trials, not as validated predictions of measured trial-level
variance. Direct validation against held-out repeated measurements is
identified in Section~\ref{sec:limitations} as required future work.

\subsection{Supplemental Synthetic Metrics}
After generation, two supplemental steps were applied:
\begin{enumerate}
\item Average power for each synthetic sample was recomputed from synthetic
energy and latency as $P = E \cdot 1000 / L$.
\item \texttt{energy\_std} was recomputed as the per-model standard deviation
of synthetic energy across repeated generated trials, after the within-model
variance was rescaled to match the grounded Apple-only seed distribution. This
calibration preserves the GAN's learned means and ranking while preventing the
small-sample WGAN from shrinking repeated-trial spread.
\end{enumerate}

\subsection{Code and Data Availability}
The benchmark scripts, the paper-aligned synthetic dataset, GAN configuration,
random seeds, and exact PyTorch and torchvision versions are available from
the author upon request. A public repository is planned at camera-ready.
Reproducibility details are provided in Appendix~\ref{sec:appendix}.

\section{Results}

\subsection{Measured Energy Ranking}
The measured ranking is shown in Fig.~\ref{fig:measured-energy}.
MobileNetV3-Small is the most energy-efficient model. EfficientNet-B0 is the
least, despite having substantially fewer FLOPs than ResNet-18.

\begin{figure}[ht]
\centering
\begin{tikzpicture}
\begin{axis}[
    ybar,
    bar width=11pt,
    width=0.95\linewidth,
    height=5.5cm,
    ylabel={Energy (J)},
    symbolic x coords={MNV3-S,MNV2,RN-18,TViT-5M,ENB0},
    xtick=data,
    x tick label style={font=\scriptsize},
    ymin=0,
    nodes near coords,
    every node near coord/.append style={font=\scriptsize},
    grid=major,
]
\addplot[fill=blue!45] coordinates {
    (MNV3-S,0.080)
    (MNV2,0.106)
    (RN-18,0.110)
    (TViT-5M,0.219)
    (ENB0,0.292)
};
\end{axis}
\end{tikzpicture}
\caption{Measured energy per inference. MNV3-S = MobileNetV3-Small, MNV2 = MobileNetV2, RN-18 = ResNet-18, TViT-5M = Tiny-ViT-5M, ENB0 = EfficientNet-B0.}
\label{fig:measured-energy}
\end{figure}

\subsection{Latency Predicts Energy; FLOPs Does Not}
\label{sec:predictors}
We fit two univariate linear models to the measured data, predicting energy
from FLOPs and from latency respectively. Results are summarized in
Table~\ref{tab:predictors}. Latency explains nearly all variance in measured
energy ($R^2 = 0.998$), while FLOPs explains only $R^2 = 0.62$.

\begin{table}[ht]
\centering
\caption{Univariate linear predictors of measured energy ($n = 5$).}
\label{tab:predictors}
\begin{tabular}{l c c}
\toprule
Predictor & $R^2$ & Pearson $r$ \\
\midrule
Latency (ms) & 0.998 & 0.999 \\
FLOPs (G) & 0.621 & 0.788 \\
Parameters (M) & 0.318 & 0.564 \\
\bottomrule
\end{tabular}
\end{table}

The contrast is sharpest for two pairs:
\begin{itemize}
\item \textbf{ResNet-18 vs.\ EfficientNet-B0:} ResNet-18 has 4.7$\times$ the
FLOPs and 2.2$\times$ the parameters of EfficientNet-B0, yet consumes
0.38$\times$ the energy. A FLOP-based predictor would rank these in the
opposite order from measurement.
\item \textbf{MobileNetV2 vs.\ ResNet-18:} ResNet-18 has 6$\times$ the FLOPs
of MobileNetV2 but only 1.04$\times$ the energy. FLOPs overstates the gap by a
factor of nearly six.
\end{itemize}

\begin{figure}[ht]
\centering
\begin{tikzpicture}
\begin{axis}[
    width=0.90\linewidth,
    height=5.5cm,
    xlabel={Latency (ms)},
    ylabel={Energy (J)},
    ymin=0,
    xmin=0,
    grid=major,
]
\addplot[
    only marks,
    mark=*,
    mark size=2.4pt,
    color=red!70!black
] coordinates {
    (15.14,0.080)
    (20.07,0.106)
    (20.68,0.110)
    (41.33,0.219)
    (55.08,0.292)
};
\end{axis}
\end{tikzpicture}
\caption{Measured energy versus latency. Linear fit gives $R^2 = 0.998$.}
\label{fig:energy-vs-latency}
\end{figure}

\subsection{Architectural Interpretation}
The latency-FLOPs decoupling is consistent with three architecture-specific
effects on a CPU memory hierarchy.

\textbf{Depthwise convolutions are FLOP-efficient but memory-bound.} The
MobileNet family relies on depthwise separable convolutions, which reduce
arithmetic intensity (FLOPs per byte of memory traffic). On a CPU without
specialized depthwise kernels, the runtime spends a higher fraction of its
time in memory-bound stalls relative to compute, so wall-clock latency drops
less than FLOPs would suggest. This explains why MobileNetV2 (0.30 G FLOPs)
runs at similar latency to MobileNetV3-Small (0.22 G FLOPs) despite the FLOP
gap.

\textbf{Compound scaling creates cache pressure.} EfficientNet-B0 uses
compound scaling that grows depth, width, and input resolution together. While
this preserves a favorable accuracy-FLOPs tradeoff, the wider activation
tensors and deeper computational graph push working sets out of L2 cache on
the CPU, generating last-level cache misses and main-memory traffic that do
not appear in the FLOP count. EfficientNet-B0's 0.39 G FLOPs corresponds to
55 ms of latency, while ResNet-18's 1.82 G FLOPs corresponds to only 21 ms.

\textbf{Transformer attention is FLOP-light but cache-unfriendly.} Tiny-ViT-5M
has fewer FLOPs than ResNet-18 (1.30 G vs.\ 1.82 G) but nearly twice the
latency (41 ms vs.\ 21 ms). The attention operation requires multiple
matrix-multiply passes over the same input tokens with different projection
weights, with limited spatial locality compared to convolutions over feature
maps. On a CPU without fused attention kernels, this becomes the latency
bottleneck.

These interpretations are not novel as systems observations
\cite{eyeriss2017,hardwareaware2019}, but they have not been previously
reported for this specific set of architectures on Apple Silicon under a
unified runtime, and they explain why latency dominates FLOPs as an energy
predictor on this platform.

\subsection{Energy-Delay Product}
Table~\ref{tab:edp} reports EDP for the five architectures. EDP penalizes
both slow and energy-hungry models. Under EDP, MobileNetV3-Small is the most
efficient model in the set, with an EDP 13.4$\times$ lower than
EfficientNet-B0.

\begin{table}[ht]
\centering
\caption{Energy-Delay Product (EDP), lower is better.}
\label{tab:edp}
\begin{tabular}{l S[table-format=2.3] S[table-format=2.2]}
\toprule
Model & {EDP (J$\cdot$ms)} & {EDP relative to MNV3-S} \\
\midrule
MobileNetV3-Small & 1.211 & 1.00 \\
MobileNetV2 & 2.127 & 1.76 \\
ResNet-18 & 2.275 & 1.88 \\
Tiny-ViT-5M & 9.051 & 7.47 \\
EfficientNet-B0 & 16.083 & 13.28 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Paper-Aligned Synthetic Fidelity}
\label{sec:fidelity}
The GAN was evaluated against the Apple-only grounded dataset.
Table~\ref{tab:fidelity} summarizes fidelity. Direct distributional tests are
applied only to the modeled quantities, energy and latency. Power is evaluated
as a derived quantity. Repeated-trial energy spread is evaluated as a
per-model summary after variance calibration. As discussed in the validation
scope subsection, this fidelity is measured against the grounded training
distribution, not against held-out measured trials.

\begin{table}[ht]
\centering
\caption{GAN fidelity relative to the grounded Apple-only benchmark.}
\label{tab:fidelity}
\small
\begin{tabular}{l c c c c}
\toprule
Feature & W & KS & $p$ & Pass \\
\midrule
Energy & 0.0039 & 0.064 & 0.156 & Yes \\
Latency & 0.6934 & 0.067 & 0.120 & Yes \\
Derived power & 0.0457 & 0.047 & 0.498 & Yes \\
Energy spread & 0.0000 & N/A & N/A & Matched \\
\bottomrule
\end{tabular}

\vspace{0.3em}
\footnotesize
W = Wasserstein distance, KS = Kolmogorov--Smirnov statistic, $p$ = KS $p$-value.
Pass criterion is $p > 0.05$.
\end{table}

\subsection{Measured versus Synthetic Means}
Table~\ref{tab:alignment} compares measured means against the synthetic means.
Energy errors range from 1.0\% to 2.3\%; latency errors range from 1.4\% to
3.9\%.

\begin{table*}[ht]
\centering
\caption{Measured versus paper-aligned synthetic mean metrics.}
\label{tab:alignment}
\begin{tabular}{l
S[table-format=1.3]
S[table-format=1.5]
S[table-format=2.2]
S[table-format=2.2]
S[table-format=2.2]
S[table-format=2.2]}
\toprule
Model &
{Measured $E$ (J)} &
{Synthetic $E$ (J)} &
{$E$ err (\%)} &
{Measured $L$ (ms)} &
{Synthetic $L$ (ms)} &
{$L$ err (\%)} \\
\midrule
MobileNetV3-Small & 0.080 & 0.08150 & 1.87 & 15.14 & 15.49 & 2.33 \\
MobileNetV2 & 0.106 & 0.10708 & 1.02 & 20.07 & 20.85 & 3.90 \\
ResNet-18 & 0.110 & 0.11169 & 1.54 & 20.68 & 20.97 & 1.39 \\
Tiny-ViT-5M & 0.219 & 0.22401 & 2.29 & 41.33 & 42.51 & 2.85 \\
EfficientNet-B0 & 0.292 & 0.29831 & 2.16 & 55.08 & 55.92 & 1.53 \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Supplemental Power and Synthetic Trial Spread}
Table~\ref{tab:supplemental} reports derived average power and synthetic
energy spread. Derived power values come directly from measured energy and
latency. Synthetic spread values come from the calibrated GAN and are
interpreted as support estimates for repeated trials.

\begin{table}[ht]
\centering
\caption{Supplemental power and synthetic trial-spread estimates.}
\label{tab:supplemental}
\small
\begin{tabular}{l
S[table-format=1.4]
S[table-format=1.4]
S[table-format=2.2]
S[table-format=1.4]}
\toprule
Model &
{$P$ (W)} &
{Syn $P$ (W)} &
{Err (\%)} &
{Syn $\sigma_E$ (J)} \\
\midrule
MobileNetV3-Small & 5.2840 & 5.2976 & 0.26 & 0.0045 \\
MobileNetV2 & 5.2815 & 5.1677 & -2.15 & 0.0081 \\
ResNet-18 & 5.3191 & 5.3558 & 0.69 & 0.0073 \\
Tiny-ViT-5M & 5.2988 & 5.3054 & 0.12 & 0.0167 \\
EfficientNet-B0 & 5.3014 & 5.3752 & 1.39 & 0.0201 \\
\bottomrule
\end{tabular}
\end{table}

\section{Discussion}
The empirical results establish that on Apple Silicon under CPU-only PyTorch
inference, latency is a near-perfect predictor of energy ($R^2 = 0.998$),
while FLOPs is substantially weaker ($R^2 = 0.62$). The architectural
interpretation in Section~\ref{sec:predictors} explains why: depthwise
convolutions, compound scaling, and transformer attention each create memory
behavior on a general-purpose CPU that decouples FLOPs from wall-clock time.
Average power is roughly constant across the five models (5.28--5.32 W), so
the model that takes longer simply pays for more wall-clock energy at a
similar power level. This means EDP, which weights both factors, ranks the
models in the same order as energy alone but with a wider efficiency spread
(13.4$\times$ best-to-worst).

The synthetic pipeline produced a tightly scoped repeated-trial model that
reproduces the measured energy ranking and tracks measured means within
1.0--2.3\% on energy and 1.4--3.9\% on latency. The strongest claim about
the synthetic dataset is that it preserves the model ranking and reproduces
mean energy and latency closely. We do not claim that the synthetic spread
estimates predict the variance of physical repeated trials, because the GAN
was trained and evaluated against a grounded distribution rather than
independently held-out measurements.

The methodological lesson from GAN development generalizes beyond this paper:
when training data are sparse, a synthetic model must be tightly scoped to
the target study and trained on normalized residual structure rather than
raw heterogeneous values. We observed this directly: a multi-device
raw-scale configuration mode-collapsed, while the same architecture trained
on combo-normalized residuals within a single-device scope reproduced
per-model structure faithfully.

\section{Limitations and Future Work}
\label{sec:limitations}
\textbf{Trial-level measurement.} The benchmark recorded means but did not
preserve per-trial spreads at sufficient resolution. Re-running each model
with a controlled number of repeated trials and reporting measured standard
deviations is the most important extension and would also enable direct
validation of the synthetic spread estimates against held-out measurements.

\textbf{Single-device scope.} The paper is a single-device study. Replicating
the benchmark on additional devices, with the same input shape, runtime
constraints, and reporting format, would test whether the FLOPs--energy
mismatch generalizes beyond Apple Silicon and would also enable a true
multi-device synthetic model.

\textbf{Power sampling resolution.} The 1 Hz \texttt{powermetrics} sampling
interval is coarse relative to per-inference latency. We addressed this with
window-level energy attribution, but higher-resolution power traces would
support per-inference rather than per-window analysis.

\textbf{Synthetic validation.} GAN fidelity is evaluated against grounded
training data. Direct validation against an independent set of measured
repeated trials is required before synthetic spreads can be cited as
predictions rather than support estimates.

\textbf{Sample size.} With $n = 5$ models, the predictor $R^2$ values in
Table~\ref{tab:predictors} are estimated from a small sample. The qualitative
gap between latency and FLOPs as predictors is large enough to be informative,
but precise $R^2$ values should be interpreted with this caveat.

\textbf{Hardware accelerator coverage.} The benchmark intentionally restricts
to CPU-only execution to control runtime variability. Apple Silicon's Neural
Engine and Metal Performance Shaders backends would likely produce different
energy rankings, which is itself a topic for future work.

\section{Conclusion}
This study finds that on Apple Silicon under CPU-only PyTorch inference,
energy efficiency cannot be reduced to FLOPs or parameter count alone.
Latency predicts energy at $R^2 = 0.998$, while FLOPs achieves only
$R^2 = 0.62$. Under Energy-Delay Product, MobileNetV3-Small is 13.4$\times$
more efficient than EfficientNet-B0, with the rest of the architectures in
between. A residualized conditional WGAN-GP, trained within a single-device
scope after an initial raw-scale multi-device configuration exhibited mode
collapse, reproduces the measured ranking and means within a few percent.
The methodological lesson is that synthetic models of hardware behavior
under sparse data must be tightly scoped and trained on normalized residual
structure rather than raw heterogeneous values.

\appendices
\section{Reproducibility}
\label{sec:appendix}

\subsection{GAN Architecture and Training}
The conditional WGAN-GP uses a generator with three fully connected layers
(64, 128, 64 units, LeakyReLU activations) and a critic with three fully
connected layers (128, 128, 1, LeakyReLU activations). Conditioning on model
identity is performed via embedding concatenation at the input layer. The
gradient penalty coefficient is $\lambda = 10$. Training uses the Adam
optimizer with $\beta_1 = 0.5$, $\beta_2 = 0.9$, and learning rate $10^{-4}$,
with five critic updates per generator update. Training proceeds for 5{,}000
generator iterations with a batch size of 64.

\subsection{Combo-Normalization Procedure}
For each (device, model) combination $c$, mean $\mu_c$ and standard deviation
$\sigma_c$ are computed for each continuous feature. Training samples are
transformed as $\tilde{x} = \mathrm{clip}((x - \mu_c)/\sigma_c, -3, 3)$.
Generated samples are inverse-transformed as $x = \mu_c + \sigma_c \cdot \tilde{x}$
using the combo identity sampled with the conditioning vector.

\subsection{Variance Calibration}
After generation, per-model variance for energy and latency is rescaled to
match the grounded Apple-only seed distribution variance. Means are unaffected
by this rescaling.

\subsection{Random Seeds and Software Versions}
Exact PyTorch and torchvision versions, NumPy and SciPy versions, random
seeds for both benchmark and GAN, and the precise window sizes used for energy
attribution are logged in the accompanying repository at camera-ready.

\begin{thebibliography}{16}

\bibitem{horowitz2014}
M.~Horowitz, ``Computing's energy problem (and what we can do about it),''
\emph{IEEE Micro}, vol.~34, no.~5, pp.~10--14, 2014.

\bibitem{mobilenet2017}
A.~G. Howard et al., ``MobileNets: Efficient convolutional neural networks for
mobile vision applications,'' \emph{arXiv preprint arXiv:1704.04861}, 2017.

\bibitem{mobilenetv22018}
M.~Sandler, A.~Howard, M.~Zhu, A.~Zhmoginov, and L.-C. Chen,
``MobileNetV2: Inverted residuals and linear bottlenecks,'' in
\emph{Proc. IEEE Conf. Computer Vision and Pattern Recognition}, 2018.

\bibitem{mobilenetv32019}
A.~Howard, M.~Sandler, G.~Chu, et al., ``Searching for MobileNetV3,'' in
\emph{Proc. IEEE International Conf. Computer Vision}, 2019.

\bibitem{efficientnet2019}
M.~Tan and Q.~Le, ``EfficientNet: Rethinking model scaling for convolutional
neural networks,'' in \emph{Proc. International Conf. Machine Learning}, 2019.

\bibitem{mittal2019}
S.~Mittal, ``A survey of techniques for improving energy efficiency of deep
learning,'' \emph{ACM Computing Surveys}, vol.~51, no.~2, 2019.

\bibitem{hardwareaware2019}
Y.~He et al., ``Understanding FLOPs is not enough: A study of hardware-aware
neural networks,'' \emph{arXiv preprint arXiv:1912.01090}, 2019.

\bibitem{eyeriss2017}
Y.-H. Chen et al., ``Eyeriss: An energy-efficient reconfigurable accelerator
for deep convolutional neural networks,'' \emph{IEEE J. Solid-State Circuits},
2017.

\end{thebibliography}

\end{document}
